Machine Learning, Building Styles and Property Values

### 2019-09-13 text/manuscript_assa.tex
vintage/era replaced by style

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Machine Learning, Building Vintage and Property Values},
            pdfauthor={Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson (University of Alabama)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Machine Learning, Building Vintage and Property Values}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson
(University of Alabama)}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{04 September 2019}

@@

\begin{document}
\maketitle
\begin{abstract}
This paper first introduces an algorithm that collects pictures of
individual buildings from Google Street View. Second, it trains a deep
convolutional neural network (CNN) to classify residential buildings
into architectural styles, taking into account spatial dependencies of
building vintages. Third, it investigates whether architectural styles
influence house prices. For re-sales, the architectural style is a
significant determinant of transaction prices while no such effect is
found for new buildings. Additionally, we are able to provide guidance
on how to detect and overcome some of the limitations of machine
learning methods through a large-scale comparison of predictions and
expert classifications.
\end{abstract}

@@

\footnote{{*Corresponding author: Thies Lindenthal (htl24@cam.ac.uk). Replication files are available from the author's Github repository \href{https://github.com/thies}{https://github.com/thies}. Paul E. Glade and Lukas Heckmann-Umhau are thanked for exceptional research assistance. Carolin Schmidt, Mike Langen, Colin Lizieri, Franz Fuerst, Stanimira Milcheva, D'Maris Coffman, Norm Miller, Stephen Malpezzi, John Clapp, Pavel Krivenko and participants at the ASSA/AREUEA 2019 annual meeting and seminars at UCL/Bartlett and the Homer Hoyt Institute offered excellent feedback and critique on earlier versions of this paper.}}

@@

If new developments were only pleasing to the eye, so Britain's
\emph{Building Better, Building Beautiful Commission}, nimbyism would
cease and housing supply could finally reach the levels demanded by a
growing and more affluent population (The Economist 2018). The inaugural
chair of this new governmental commission boldly suggests that we should
\emph{``build, as our Georgian and Victorian forebears built}
{[}\ldots{}{]}. \emph{All objections to new building would slip away in
the sheer relief of the public''} (Scruton 2018). Even Prince Charles,
in similar spirit, put forward ten principles for urban growth and
architecture that emphasize tradition and aesthetics (HRH The Prince of
Wales 2014). When governments and princes occupy themselves with beauty,
economists surely may take a closer look as well. In this paper, we aim
to identify the architectural styles of residential buildings using
computer vision techniques and to empirically search for any transaction
price premia associated with these styles. If aesthetics were as strong
a force in the built environment as claimed, we should indeed find a
significant effect on property prices.

This paper assesses the value of architectural aesthetics heads on by
classifying the vintage of residential buildings using street level
images. Recent work illustrates how street level imagery and machine
learning classification is an efficient and powerful combination for
measuring previously unobserved characteristics of the urban
environment. Naik et al. (2017) describe how neighborhood demographics
may impact the physical appearance of neighborhoods. Gebru et al. (2017)
use classified vehicle make and model to predict income, race,
education, and voting patterns at the precinct level. Glaeser et al.
(2018) predict income in New York City. Naik, Raskar, and Hidalgo (2016)
create a neighborhood safety based Streetscore which is shown to be
highly correlated with neighborhood population density and household
income. De Nadai et al. (2016) find that greenery and street facing
windows contribute to a positive appearance of safety while Liu et al.
(2017) evaluate the quality and upkeep of the built environment along
Beijing's streets.

@@

In this paper we suggest a method to collect a large number of images of
individual UK buildings from Google Street View, classify the depicted
buildings using deep convolutional neural networks, combine the derived
information with sales price data and, ultimately, estimate marginal
prices for the estimated building characteristics. Our work demonstrates
that utilizing pretrained convolutional neural networks to detect
complex characteristics such as housing vintages with relatively high
accuracy comes at low computational costs and is feasible with only
modestly-sized training data. Additionally, we illustrate how machine
classifications may be sensitive to photo quality by comparing predicted
vintages to those made by human experts.

\hypertarget{building-vintage-and-property-values}{%
\subsection{Building vintage and property
values}\label{building-vintage-and-property-values}}

The aesthetics of different architectural
vintages\footnote{We use ``style'',``era'' or ``vintage'' interchangeably in this paper.}
make for an easy cocktail party conversation topic -- and they may also
influence market prices. Using architectural assessments by human
experts, Buitelaar and Schilder (2017) find a sizable premium of 5\% for
new buildings in the Netherlands that refer to traditional styles and a
staggering 15\% premium for new buildings that closely follow
traditional shapes, facade composition and details. Their study
carefully disentangles the architectural style from other unobserved
characteristics such as building quality, differences in location or
year of construction. These controls are crucial, as earlier work has
established that age and vintage variables tend to be highly correlated.
Coulson and McMillen (2008), for instance, suggest a non-parametric
estimator and establish a U-shaped age function and distinct price
discounts for postwar and contemporary vintages (vis-à-vis more historic
vintages). Francke and Minne (2017) investigate the depreciation of
residential real estate in the Netherlands and decompose land versus
structure values singling out the effect of ``physical deterioration,
functional obsolescence and and vintage effects''. They find that
buildings from the 1930s carry a strong price premium.

@@

Glaeser, Kincaid, and Naik (2018) rely on an different ILSVRC
competitor, \emph{Resnet-101} (He et al. 2016). They reduce the
extracted feature vectors to lower dimensionality (1024 to 100
dimensions) based on principal component analysis (PCA). We follow a
different strategy and double the dimensionality by including the
feature vector of the building's nearest neighbor. Doubling up allows us
to model spatial dependencies in building styles, similar to spatially
correlated land cover classifications in Ghimire, Rogan, and Miller
(2010).

@@

Cambridge's houses can be classified into seven broad eras:
\footnote{We are grateful to colleagues from the Architecture Department at the University of Cambridge to provide these general vintage descriptions.}

\begin{itemize}
\item
  \emph{Georgian} (c1714--1837) houses feature key characteristics such
  as sash windows, fan lights above doors, the use of stucco on facades,
  often wrought work grilles, railings etc.\label{def_eras}
\item
  In the \emph{Early Victorian} era (c1837--c1870s), a growing taste for
  individualized embellishment led to the development of elaborate
  features such as carved barge boards or finials. The development of
  sheet glass led to sash windows becoming more affordable, and,
  increasingly, wider.
\item
  In the \emph{Late Victorian} era (c1870s--1901), bay windows became
  more and more widespread, and increasingly substantial. Stylistic
  movements such as the Queen Anne revival style contributed richly
  ornamental details to the formal repertoire employed by designers.
  Stained glass became more popular. \emph{Edwardian} architecture
  (1901-1910) tends to be less ornate than late Victorian architecture.
\item
  The \emph{Interwar} period (1918--1939) saw the cost of building
  construction fall, amidst a drive to provide better housing for the
  working classes. New housing types were being favored.
\item
  The \emph{Postwar} (1950-1980) era continued on this path, with an
  embrace of high-rise as well as low rise housing. Facades vary greatly
  between brick, tiling, pebbledash and render.
\item
  The cut-off year for our \emph{Contemporary} era to begin is 1980.
  \emph{Revival} are contemporary buildings trying to emulate historic,
  mostly Victorian, architecture.
\end{itemize}

Two final-year architectural students classified a large sub-sample of
approximately 25,000 images from our data set of Cambridge houses. This
is a much larger sample than needed in our case. In our case, each
category requires less than 250 observations to reach saturated training
accuracy levels. We greatly exceed this number for the purpose of this
paper so that we can compare the out-of-sample convolutional neural
network predictions to the ground truth as assigned by the experts. This
allows us to examine the power and size of the assignment tests. In
addition having both human and machine classification for a large sample
of the data allows for robustness checks on the machine comparisons.

We create stratified training samples of 600 buildings from each
category but \emph{Georgian}, which is the least common style in
Cambridge and for which we can only sample 330 examples. This leaves a
sufficiently sized out-of-sample verification data set.

@@

We tabulate confusion matrices and evaluate the classification
performance in terms of recall, precision and \(F_1\)-scores. The next
step regresses the true vintage observed, captured by binomial variables
\(D_{TrueVint}\), against a vector of hedonic variables \(\mathbf{X}\),
vectors of year \(\mathbf{Y}\) and neighborhood \(\bm{Loc}\) dummy
variables and the building's \(\mathbf{Vintage}\), estimated both with
and without spatial dependencies. The intercept is denoted by \(\alpha\)
while \(\bm{\beta}\), \(\bm{\delta}\), \(\bm{\gamma}\) and
\(\bm{\lambda}\) are vectors of regression coefficients, and
\(\epsilon\) is the error term.

\begin{equation} \label{eq:rev}
  logit( D_{TrueVint,i} ) = \alpha + \beta\mathbf{X_i} + \delta\mathbf{Vintage}_i +  \gamma\mathbf{Y}_i + \lambda\mathbf{Loc}_i +  \epsilon_i
\end{equation}

@@

Will the resulting classifiers be biased with respect to property
values, picking up correlated but irrelevant cues such as greenery,
upkeep, or cars brands instead of building exteriors? Are, for instance,
more valuable houses more likely to be classified as the popular
\emph{Late Victorian} vintage while starter homes are disproportionally
more likely to be classified as e.g. \emph{Early Victorian}? We expect
that in cases in which the ensemble provided a high confidence score,
building features are clearly detectable on the pictures. For close
calls, however, where models within the ensemble cannot agree, other
factors might tilt the scale towards an incorrect class. A sizable
difference in estimated marginal prices for estimated vintages versus
values for ground truth vintages in the subsequent hedonic regression
(Eq. \ref{eq:hed}, below) would indicate such a bias.

\hypertarget{building-era-and-property-price}{%
\subsection{Building era and property
price}\label{building-era-and-property-price}}

Merging the estimated vintage classification with sales data (Land
Registry 2017) we estimate a hedonic regression equation that
establishes marginal prices for the building vintage (similar to
Moorhouse and Smith 1994; Asabere, Hachey, and Grubaugh 1989; Vandell
and Lane 1989; Fuerst, McAllister, and Murray 2011; Plaut and Uzulena
2006), among other characteristics: \begin{equation} \label{eq:hed}
  ln(Price_i) = \alpha + \beta\mathbf{X_i} + \delta\mathbf{Vintage}_i + \eta\mathbf{VintNeigh}_i + \iota\mathbf{Vintage\cdot VintNeigh}_i  +\gamma\mathbf{Y}_i + \lambda\mathbf{Loc}_i +  \epsilon_i
\end{equation}

Here, the natural logarithm of sales prices is explained by a linear
combination of hedonic attributes described in vector \(\mathbf{X}\),
vectors of year \(\mathbf{Y}\) and neighborhood \(\bm{Loc}\) dummy
variables and the building's estimated \(\mathbf{Vintage}\) and the
prevailing vintage of other buildings in the direct proximity
\((\mathbf{VintNeigh})\). The intercept is denoted by \(\alpha\) while
\(\bm{\beta}\), \(\bm{\delta}\), \(\bm{\eta}\), \(\bm{\iota}\),
\(\bm{\gamma}\) and \(\bm{\lambda}\) are vectors of regression
coefficients. \(\epsilon\) is the IID error term. Heteroscedastic robust
standard errors will be reported.

Are buildings from different eras imperfect substitutes catering to
multiple groups of households with distinct vintage preferences?
Buitelaar and Schilder (2017) indicate that any premium for an
architectural style must stem from either differences in construction
prices (which they do not find in their Dutch sample) or from supply
constraints, as new construction potentially does not capture the demand
for traditional styles. For Cambridge, new supply will inevitably be of
either contemporary or revival style as historic vintage buildings, by
definition, are not supplied any more. Estimating Eq. \ref{eq:hed} for a
subset of newly constructed buildings will show whether construction
prices or supply constraints for new homes built according to different
architectural styles persist -- if too few vernacular buildings were
built, prices should reflect such a shortage.

@@

Overall, we collect pictures of more than 48,000 distinct buildings from
the Google Street View API. The subset of 25,000 pictures that has been
classified by trained architects into the seven vintages will serve as
\emph{ground truth} for our study. Two additional categories identify
buildings hidden by \emph{greenery} or uninformative pictures featuring
\emph{plain walls}. The classified images are available at the authors'
websites.

@@

Almost 16,000 of the 25,000 buildings for which we have the architects'
ground truth could be matched to the 23,768 sales transactions that have
been recorded for Cambridge between 1995 and 2018. Thus, we observe both
the machine \emph{estimated vintage} and architect defined \emph{ground
truth} classification for the majority of buildings in the data set.

@@

The agreement between vintage estimates and the corresponding ground
truth classifications by experts is reassuringly strong. When relying on
feature vectors of the buildings and their closest neighbors, 67 to 84
percent of true classifications are matched with correctly estimated
labels (Table \ref{tab:confmat}, Panel A). The recall rates are
especially high for older vintages: For \emph{Georgians}, it is 79
percent, \emph{Early Victorian} 81 percent, \emph{Late
Victorian/Edwardian} 78 percent, \emph{Interwar} 84 percent,
respectively. For more recent \emph{Postwar} buildings, the rate drops
to 71 percent, for \emph{Contemporary} to 72 percent and for
\emph{Revival} to 67 percent.

A visual inspection of classifications is additionally reassuring.
Figure \ref{fig:vint_examples} displays the pictures which carry the
highest out-of-sample scores for each vintage. The model is clearly able
to differentiate based on small cues, even when only parts of the facade
are captured on the picture.

The Herfindahl scores for classifications vary substantially across
ground truth vintages (Table \ref{tab:herftab}), indicating that the
models within the ensemble, for instance, agree more often on a building
being \emph{Georgian} than finding consensus for other vintages. For
\emph{Contemporary} or \emph{Revival} vintages, individual
classifications differ most. All off-diagonal elements in the lower
panel of Table \ref{tab:herftab} are negative, suggesting that, in
general, the Herfindahl index is a good predictor of misclassified
images.

@@

The number of misclassifications decreases decidedly when considering
observations with high ensemble confidence only: The majority vote based
on the most confident two-thirds of predictions only are more in line
with ground truth, as higher recall, precision and \(F_1\)-scores in
Table \ref{tab:confmat}, Panel B, show. The improvement is most evident
in the precision rates. For the relatively rare categories of
\emph{Georgian} and \emph{Revival} about the same number of buildings
classified as \emph{Georgian/Revival} in Panel A actually belong to a
different ground truth class -- the precision shoots up from 0.50 to
0.70 for \emph{Georgian} and from 0.52 to 0.82 for \emph{Revival} when
excluding the not so confident lower third of predictions in Panel B.

@@

Overall, the spatial softmax classification incorporating the nearest
neighbor's feature vector outperforms the base classifier: pairwise
differences of the \(F_1\)-scores from the spatial and the base
classifier are clearly positive (Fig. \ref{fig:boxplot}). AIC values for
28 binomial models based on Eq. \ref{eq:rev} decrease as hedonic
characteristics, location and time dummies are augmented by the base and
spatial ML predictor. For all building vintages, the AIC are lowest for
the spatial ML predictor. In sum, the spatial classifier performs better
for UK vintage data than a classifier based on single buildings only.

@@

Second, we developed a new database of 25,000 building pictures that
have been classified by architecture experts into relevant architectural
styles or vintages. We subsequently trained a neural network classifier
to automatically classify all residential buildings of a mid-sized
English city into architectural styles or vintages. The suggested
classifier is trained on feature vectors of buildings and their nearest
neighbors to exploit spatial correlation in observed classifications.

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Machine Learning, Building Styles and Property Values},
            pdfauthor={Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson (University of Alabama)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Machine Learning, Building Styles and Property Values}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson
(University of Alabama)}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{13 September 2019}

@@

\begin{document}
\maketitle
\begin{abstract}
This paper first introduces an algorithm that collects pictures of
individual buildings from Google Street View. Second, it trains a deep
convolutional neural network (CNN) to classify residential buildings
into architectural styles, taking into account spatial dependencies of
these styles. Third, it investigates whether architectural styles
influence house prices. For re-sales, the architectural style is a
significant determinant of transaction prices while no such effect is
found for new buildings. Additionally, we are able to provide guidance
on how to detect and overcome some of the limitations of machine
learning methods through a large-scale comparison of predictions and
expert classifications.
\end{abstract}

@@

\footnote{{*Corresponding author: Thies Lindenthal (htl24@cam.ac.uk). Replication files are available from the author's Github repository \href{https://github.com/thies}{https://github.com/thies}. Paul E. Glade and Lukas Heckmann-Umhau are thanked for exceptional research assistance. Carolin Schmidt, Peter Brummund, Mike Langen, Colin Lizieri, Franz Fuerst, Stanimira Milcheva, D'Maris Coffman, Norm Miller, Stephen Malpezzi, John Clapp, Pavel Krivenko and participants at the ASSA/AREUEA 2019 annual meeting and seminars at UCL/Bartlett and the Homer Hoyt Institute offered excellent feedback and critique on earlier versions of this paper.}}

@@

If new developments were only pleasing to the eye, so Britain's
\emph{Building Better, Building Beautiful Commission} advising
Government, nimbyism would cease and housing supply could finally reach
the levels demanded by a growing and more affluent population (The
Economist 2018). The inaugural chair of this new commission boldly
suggested that if we built \emph{``as our Georgian and Victorian
forebears built} {[}\ldots{}{]}. \emph{All objections to new building
would slip away in the sheer relief of the public''} (Scruton 2018).
Even Prince Charles, in similar spirit, put forward ten principles for
urban growth and architecture that emphasize tradition and aesthetics
(HRH The Prince of Wales 2014). When governments and princes occupy
themselves with beauty, economists surely may take a closer look as
well. In this paper, we aim to identify the architectural styles of
residential buildings using computer vision techniques and to
empirically search for any transaction price premia associated with
these styles. If aesthetics were as strong a force in the built
environment as claimed, we should indeed find a significant effect on
property prices.

This paper assesses the value of architectural aesthetics head on by
classifying the exterior styles of residential buildings using street
level images. Recent work illustrates how street level imagery and
machine learning classification is an efficient and powerful combination
for measuring previously unobserved characteristics of the urban
environment. Naik et al. (2017) describe how neighborhood demographics
may impact the physical appearance of neighborhoods. Gebru et al. (2017)
use classified vehicle make and model information to predict income,
race, education, and voting patterns at the precinct level. Glaeser et
al. (2018) predict income in New York City. Naik, Raskar, and Hidalgo
(2016) create a neighborhood safety based Streetscore which is shown to
be highly correlated with neighborhood population density and household
income. De Nadai et al. (2016) find that greenery and street facing
windows contribute to a positive appearance of safety while Liu et al.
(2017) evaluate the quality and upkeep of the built environment along
Beijing's streets.

@@

In this paper we suggest a method to collect a large number of images of
individual UK buildings from Google Street View, classify the depicted
buildings using deep convolutional neural networks, combine the derived
information with sales price data and, ultimately, estimate marginal
prices for the estimated building characteristics. Our work demonstrates
that utilizing pretrained convolutional neural networks to detect
complex characteristics such as housing appearances with relatively high
accuracy comes at low computational costs and is feasible with only
modestly-sized training data. Additionally, we illustrate how machine
classifications may be sensitive to photo quality by comparing machine
classifications to those made by human experts.

\hypertarget{building-styles-and-property-values}{%
\subsection{Building styles and property
values}\label{building-styles-and-property-values}}

The aesthetics of different architectural styles make for an easy
cocktail party conversation topic -- and they may also influence market
prices. Using architectural assessments by human experts, Buitelaar and
Schilder (2017) find a sizable premium of 5\% for new buildings in the
Netherlands that refer to traditional styles and a staggering 15\%
premium for new buildings that closely follow traditional shapes, facade
composition and details. Their study carefully disentangles the
architectural style from other unobserved characteristics such as
building quality, differences in location, or year of construction.
These controls are crucial, as earlier work has established that age and
style variables tend to be highly correlated. Coulson and McMillen
(2008), for instance, suggest a non-parametric estimator and establish a
U-shaped age function and distinct price discounts for postwar and
contemporary styles (vis-à-vis more historic styles). Francke and Minne
(2017) investigate the depreciation of residential real estate in the
Netherlands and decompose land versus structure values singling out the
effect of ``physical deterioration, functional obsolescence and and
vintage effects''. They find that buildings from the 1930s carry a
strong price premium.

@@

Glaeser, Kincaid, and Naik (2018) rely on an different ILSVRC
competitor, \emph{Resnet-101} (He et al. 2016). They reduce the
extracted feature vectors to lower dimensionality (1024 to 100
dimensions) based on principal component analysis (PCA). We follow a
different strategy and double the dimensionality by including the
feature vector of the closest building in the sample, which, in most
cases, is the direct neighbor. The feature vectors of neighbors have
been collected in exactly the same way as those for other buildings in
the sample. Doubling up allows us to model spatial dependencies in
building styles, similar to spatially correlated land cover
classifications in Ghimire, Rogan, and Miller (2010).

@@

Cambridge's houses can be classified into seven broad styles:
\footnote{We are grateful to colleagues from the Architecture Department at the University of Cambridge to provide these general style descriptions.}

\begin{itemize}
\item
  \emph{Georgian} (c1714--1837) houses feature key characteristics such
  as sash windows, fan lights above doors, the use of stucco on facades,
  often wrought work grilles, railings etc.\label{def_eras}
\item
  In the \emph{Early Victorian} era (c1837--c1870s), a growing taste for
  individualized embellishment led to the development of elaborate
  features such as carved barge boards or finials. The development of
  sheet glass led to sash windows becoming more affordable, and,
  increasingly, wider.
\item
  In the \emph{Late Victorian} era (c1870s--1901), bay windows became
  more and more widespread, and increasingly substantial. Stylistic
  movements such as the Queen Anne revival style contributed richly
  ornamental details to the formal repertoire employed by designers.
  Stained glass became more popular.
\item
  \emph{Edwardian} architecture (1901-1910) tends to be less ornate than
  late Victorian architecture.
\item
  The \emph{Interwar} period (1918--1939) saw the cost of building
  construction fall, amidst a drive to provide better housing for the
  working classes. New housing types were being favored.
\item
  The \emph{Postwar} (1950-1980) styles continued on this path, with an
  embrace of high-rise as well as low rise housing. Facades vary greatly
  between brick, tiling, pebbledash and render.
\item
  The cut-off year for our \emph{Contemporary} era to begin is 1980.
  \emph{Revival} style-homes are contemporary buildings trying to
  emulate historic, mostly Victorian, architecture.
\end{itemize}

Two final-year architectural students classified a large sub-sample of
approximately 25,000 images from our data set of Cambridge houses. This
is a much larger sample than needed in our case. In our case, each
category requires less than 250 observations to reach saturated training
accuracy levels. We greatly exceed this number for the purpose of this
paper so that we can compare the out-of-sample convolutional neural
network predictions to the architects' classifications. This allows us
to examine the power and size of the assignment tests. In addition
having both human and machine classification for a large sample of the
data allows for robustness checks on the machine comparisons.

We create stratified training samples of 600 buildings from each
category but \emph{Georgian}, which is the least common style in
Cambridge and for which we can only sample 330 examples. This leaves a
sufficiently sized out-of-sample verification data set of 21,000 homes.

@@

We tabulate confusion matrices and evaluate the classification
performance in terms of recall, precision and \(F_1\)-scores. More
formally, the next step regresses the style classifications by
architects, captured by binomial variables \(D_{TrueVint}\), against a
vector of hedonic variables \(\mathbf{X}\), vectors of year
\(\mathbf{Y}\) and neighborhood \(\bm{Loc}\) dummy variables and the
building's \(\mathbf{Style}\), estimated both with and without spatial
dependencies. The intercept is denoted by \(\alpha\) while
\(\bm{\beta}\), \(\bm{\delta}\), \(\bm{\gamma}\) and \(\bm{\lambda}\)
are vectors of regression coefficients, and \(\epsilon\) is the error
term.

\begin{equation} \label{eq:rev}
  logit( D_{TrueVint,i} ) = \alpha + \beta\mathbf{X_i} + \delta\mathbf{Style}_i +  \gamma\mathbf{Y}_i + \lambda\mathbf{Loc}_i +  \epsilon_i
\end{equation}

@@

Will the resulting classifiers be biased with respect to property
values, picking up correlated but irrelevant cues such as greenery,
upkeep, or cars brands instead of building exteriors? Are, for instance,
more valuable houses more likely to be classified as the popular
\emph{Late Victorian} style while starter homes are disproportionally
more likely to be classified as e.g. \emph{Early Victorian}? We expect
that in cases in which the ensemble provided a high confidence score,
building features are clearly detectable on the pictures. For close
calls, however, where models within the ensemble cannot agree, other
factors might tilt the scale towards an incorrect class. A sizable
difference in estimated marginal prices for machine's classifications
versus the estimates for architects' classifications in the subsequent
hedonic regression (Eq. \ref{eq:hed}, below) would indicate such a bias.

\hypertarget{building-style-and-property-price}{%
\subsection{Building style and property
price}\label{building-style-and-property-price}}

Merging the machine's classification with sales data (Land Registry
2017) we estimate a hedonic regression equation that establishes
marginal prices for the building style (similar to Moorhouse and Smith
1994; Asabere, Hachey, and Grubaugh 1989; Vandell and Lane 1989; Fuerst,
McAllister, and Murray 2011; Plaut and Uzulena 2006), among other
characteristics: \begin{equation} \label{eq:hed}
  ln(Price_i) = \alpha + \beta\mathbf{X_i} + \delta\mathbf{Style}_i + \eta\mathbf{StyleNeigh}_i + \iota\mathbf{Style\cdot StyleNeigh}_i  +\gamma\mathbf{Y}_i + \lambda\mathbf{Loc}_i +  \epsilon_i
\end{equation}

Here, the natural logarithm of sales prices is explained by a linear
combination of hedonic attributes described in vector \(\mathbf{X}\),
vectors of year \(\mathbf{Y}\) and neighborhood \(\bm{Loc}\) dummy
variables and the building's estimated \(\mathbf{Style}\) and the
prevailing vintage of other buildings in the direct proximity
\((\mathbf{StyleNeigh})\). The intercept is denoted by \(\alpha\) while
\(\bm{\beta}\), \(\bm{\delta}\), \(\bm{\eta}\), \(\bm{\iota}\),
\(\bm{\gamma}\) and \(\bm{\lambda}\) are vectors of regression
coefficients. \(\epsilon\) is the IID error term. Heteroscedastic robust
standard errors will be reported.

Are buildings with different appearances imperfect substitutes catering
to multiple groups of households with distinct style preferences?
Buitelaar and Schilder (2017) indicate that any premium for an
architectural style must stem from either differences in construction
prices (which they do not find in their Dutch sample) or from supply
constraints, as new construction potentially does not capture the demand
for traditional styles. For Cambridge, new supply will inevitably be of
either contemporary or revival style as historic vintages (and their
styles), by definition, are not supplied any more. Estimating Eq.
\ref{eq:hed} for a subset of newly constructed buildings will show
whether construction prices or supply constraints for new homes built
according to different architectural styles persist -- if too few
vernacular buildings were built, prices should reflect such a shortage.

@@

Overall, we collect pictures of more than 48,000 distinct buildings from
the Google Street View API. The subset of 25,000 pictures that has been
classified by trained architects into the seven styles will serve as
\emph{ground truth} for our study. Two additional categories identify
buildings hidden by \emph{greenery} or uninformative pictures featuring
\emph{plain walls}. The classified images are available at the authors'
websites.

@@

Almost 16,000 of the 25,000 buildings for which we have an architect's
classification could be matched to the 23,768 sales transactions that
have been recorded for Cambridge between 1995 and 2018. The remaining
9,000 houses have not been sold since 1995 - but the information on
their architectural style is still valuable for our measures of
neighborhood styles. Thus, we observe both the machine's and the
architects' classification for the majority of buildings in the data
set.

@@

The agreement between classifications by the machine and by the
architects is reassuringly strong. When relying on feature vectors of
the buildings and their closest neighbors, 67 to 84 percent of true
classifications are matched with correctly estimated labels (Table
\ref{tab:confmat}, Panel A). The recall rates are especially high for
older styles: For \emph{Georgians}, it is 79 percent, \emph{Early
Victorian} 81 percent, \emph{Late Victorian/Edwardian} 78 percent,
\emph{Interwar} 84 percent, respectively. For more recent \emph{Postwar}
buildings, the rate drops to 71 percent, for \emph{Contemporary} to 72
percent and for \emph{Revival} to 67 percent.

A visual inspection of classifications is additionally reassuring.
Figure \ref{fig:vint_examples} displays the pictures which carry the
highest out-of-sample scores for each style The model is clearly able to
differentiate based on small cues, even when only parts of the facade
are captured on the picture.

The Herfindahl scores for classifications vary substantially across
styles, as classified by the architects (Table \ref{tab:herftab}),
indicating that the models within the ensemble, for instance, agree more
often on a building being \emph{Georgian} than finding consensus for
other styles. For \emph{Contemporary} or \emph{Revival} styles,
individual classifications differ most. All off-diagonal elements in the
lower panel of Table \ref{tab:herftab} are negative, suggesting that, in
general, the Herfindahl index is a good predictor of misclassified
images.

@@

The number of misclassifications decreases decidedly when considering
observations with high ensemble confidence only: The majority vote based
on the most confident two-thirds of machine classifications only are
more in line with the architects' classifications, as higher recall,
precision and \(F_1\)-scores in Table \ref{tab:confmat}, Panel B, show.
The improvement is most evident in the precision rates. For the
relatively rare categories of \emph{Georgian} and \emph{Revival} about
the same number of buildings classified as \emph{Georgian/Revival} in
Panel A actually belong to a different style, according to the human
experts. The precision shoots up from 0.50 to 0.70 for \emph{Georgian}
and from 0.52 to 0.82 for \emph{Revival} when excluding the not so
confident lower third of predictions in Panel B.

@@

Overall, the spatial softmax classification incorporating the nearest
neighbor's feature vector outperforms the base classifier: pairwise
differences of the \(F_1\)-scores from the spatial and the base
classifier are clearly positive (Fig. \ref{fig:boxplot}). AIC values for
28 binomial models based on Eq. \ref{eq:rev} decrease as hedonic
characteristics, location and time dummies are augmented by the base and
spatial ML predictor (Table \ref{tab:aic}). For all building styles, the
AIC are lowest for the spatial ML predictor. In sum, the spatial
classifier performs better for UK style data than a classifier based on
single buildings only.

@@

Second, we developed a new database of 25,000 building pictures that
have been classified by architecture experts into relevant architectural
styles. We subsequently trained a neural network classifier to
automatically classify all residential buildings of a mid-sized English
city into architectural styles. The suggested classifier is trained on
feature vectors of buildings and their nearest neighbors to exploit
spatial correlation in observed classifications.
